### 2024-08-05 mixed-models.qmd
20240731 - placeholder
20240805 - bayesian mixed model

```{r}
library("lme4")
library("lmerTest")
library("MuMIn")
library("emmeans")
library("sjstats")
library("mgcv")
library("AICcmodavg")
library("bbmle")
library("parallel")
library("plotly")
library("viridis")
library("tidyverse")
```

@@

>>>>

```{r}
library("lme4")
library("lmerTest")
library("MuMIn")
library("emmeans")
library("sjstats")
library("mgcv")
library("AICcmodavg")
library("bbmle")
library("rstan")
library("brms")
library("cmdstanr") # todo: install.packages("cmdstanr", repos = c('https://stan-dev.r-universe.dev', getOption("repos"))); check_cmdstan_toolchain(); cmdstanr::install_cmdstan()
library("parallel")
library("parallely")
library("plotly")
library("viridis")
library("tidyverse")
```

@@

### Bayesian Mixed Models {#sec-fantasyPointsByAgeExperienceModelsBayesian}

#### Determine Response Distribution {#sec-fantasyPointsByAgeExperienceModelsBayesianDistribution}

```{r}
#| label: fig-histogramFantasyPoints
#| fig-cap: "Histogram of Fantasy Points with Overlaid Density Plot and Rug Plot."

ggplot2::ggplot(
  data = player_stats_seasonal_offense_subset,
  mapping = aes(
    x = fantasy_points)
) +
  geom_histogram(
    aes(y = after_stat(density)),
    color = "#000000",
    fill = "#0099F8"
  ) +
  geom_density(
    color = "#000000",
    fill = "#F85700",
    alpha = 0.6 # add transparency
  ) +
  geom_rug() +
  theme_classic()
```

Because the response distribution (i.e., fantasy points) is positively skewed, we will use a gamma response distribution.

#### Prepare Data {#sec-fantasyPointsByAgeExperienceModelsBayesianPrepareData}

```{r}
player_stats_seasonal_offense_subset$fantasy_points_posOnly <- player_stats_seasonal_offense_subset$fantasy_points
player_stats_seasonal_offense_subset$fantasy_points_posOnly[which(player_stats_seasonal_offense_subset$fantasy_points_posOnly < 0)] <- 0
```

#### Specify Model Formula {#sec-fantasyPointsByAgeExperienceModelsBayesianFormula}

Specify model formula:

```{r}
#| eval: false

#mgcv smooth terms: https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/smooth.terms.html

#bayesianMixedModelFormula <- brms::bf(
#  fantasy_points ~ positionFactor + s(ageCentered20, by = #positionFactor) + years_of_experience + s(ageCentered20, #player_idFactor, bs = "fs") + (1 | player_idFactor)
#)

#bayesianMixedModelFormula <- brms::bf(
#  fantasy_points ~ positionFactor + s(ageCentered20, by = #positionFactor) + years_of_experience + (1 | player_idFactor)
#)

bayesianMixedModelFormula <- brms::bf(
  fantasy_points_posOnly ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(ageCentered20, player_idFactor, bs = "re") + (1 | player_idFactor)
)
```

#### Run Model {#sec-fantasyPointsByAgeExperienceModelsBayesianRunModel}

Now, we can run the model.

::: {#nte-bayesianMixedModel .callout-note title="Bayesian mixed model"}
Note: the following code takes a while to run.
:::

```{r}
#| eval: false

bayesianMixedModelFit <- brms::brm(
  formula = bayesianMixedModelFormula,
  data = player_stats_seasonal_offense_subset,
  family = hurdle_gamma(),
  cores = 4,
  save_pars = save_pars(latent = FALSE, all = FALSE),
  threads = threading(parallelly::availableCores()),
  backend = "cmdstanr",
  seed = 52242,
  silent = 0
)
```

save(
  bayesianMixedModelFit,
  file = "./data/bayesianMixedModelFit.RData")
```

```{r}
#| include: false

load("./data/bayesianMixedModelFit.RData")
```

#### Model Summary {#sec-fantasyPointsByAgeExperienceModelsBayesianModelSummary}

```{r}
summary(bayesianMixedModelFit)
brms::prior_summary(bayesianMixedModelFit)
```

#### Trace Plots {#sec-fantasyPointsByAgeExperienceModelsBayesianTrace}

```{r}
#| label: fig-bayesianTracePlots
#| fig-cap: "Trace Plots from Bayesian Mixed Model."

plot(bayesianMixedModelFit, ask = FALSE)
```

#### Posterior Predictive Check {#sec-fantasyPointsByAgeExperienceModelsBayesianPPcheck}

```{r}
#| label: fig-bayesianPosteriorPredictiveCheck
#| fig-cap: "Posterior Predictive Check from Bayesian Mixed Model."

pp_check(bayesianMixedModelFit) + 
  ggplot2::xlim(0, 600)
```
